Issues Developing the new Abalone MSE

### 2020-06-10 README.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SMUs), within a region:

* region - highest geogrpahical level. Is simply the totality of the spatial management units.

* SMU - spatial management units. In Tasmania these would currently be the classical statistical blocks.

@@

This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case that the general properties of each SMU are used to set the scene for each SMU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SMU.

@@

This is a basic example which illustrates the generation of an initial equilibrium. It uses built in data-sets but usually you would generate regionC (constants) and regionD (the dynamic components) from the conditioning data:

```{r example}
# a minimal example
starttime <- as.character(Sys.time())
library(aMSE)
# Obviously you should modify the resdir to suit your own computer
resdir <- "./../../rcode2/aMSEUse/out/testrun"
dirExists(resdir,make=TRUE,verbose=TRUE)
# You now need to ensure that there is a control.csv, reg1smu2pop6.csv
# and region1.csv file in the data directory
ctrlfiletemplate(resdir)  # puts a template ctrlfile into resdir
regionfiletemplate(resdir)  # puts a template region file into resdir
datafiletemplate(6,resdir,filename="reg1smu2pop6.csv") # etc
ctrl <- checkresdir(resdir) # checks the data files are present
runname <- ctrl$runlabel    # and reads the control file, if present
region1 <- readregionfile(resdir,ctrl$regionfile)
glb <- region1$globals
constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)

out <- setupregion(constants, region1) # make operating model 
regionC <- out$regionC     # constant bits
regionD <- out$regionD     # dynamics bits
product <- out$product     # all usually saved in resdir, not here
glb <- out$glb             # now contains the larval movement matrix
  # Are we at equilibrium?
regDe <- testequil(regionC,regionD,glb)

@@

plotproductivity(resdir,runname,product,glb)
biology_plots(resdir, runname, glb, regionC)
numbersatsize(resdir, runname, glb, regionD)

@@

reportlist <- list(runname=runname,
                   starttime=starttime,endtime=endtime,
                   regionC=regionC, regionD=regionD, product=product,
                   glb=glb,constants=constants
)

>>>>

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:

* zone - highest geogrpahical level. Is simply the totality of the spatial management units.

* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.

@@

This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU.

@@

This is a basic example which illustrates the generation of an initial equilibrium. It uses built in data-sets but usually you would generate zoneC (constants) and zoneD (the dynamic components) from the conditioning data:

```{r example}
# a minimal example
starttime <- as.character(Sys.time())
library(aMSE)
library(makehtml)
# Obviously you should modify the resdir to suit your own computer
resdir <- "./../../rcode2/aMSEUse/out/testrun"
dirExists(resdir,make=TRUE,verbose=TRUE)
# You now need to ensure that there is a control.csv, reg1sau2pop6.csv
# and zone1.csv file in the data directory
ctrlfiletemplate(resdir)  # puts a template ctrlfile into resdir
zonefiletemplate(resdir)  # puts a template zone file into resdir
datafiletemplate(6,resdir,filename="zone1sau2pop6.csv") # etc
ctrl <- checkresdir(resdir) # checks the data files are present
runname <- ctrl$runlabel    # and reads the control file, if present
zone1 <- readzonefile(resdir,ctrl$zonefile)
glb <- zone1$globals
constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)

out <- setupzone(constants, zone1) # make operating model 
zoneC <- out$zoneC     # constant bits
zoneD <- out$zoneD     # dynamics bits
product <- out$product     # all usually saved in resdir, not here
glb <- out$glb             # now contains the larval movement matrix
  # Are we at equilibrium?
regDe <- testequil(zoneC,zoneD,glb)

@@

plotproductivity(resdir,runname,product,glb)
biology_plots(resdir, runname, glb, zoneC)
numbersatsize(resdir, runname, glb, zoneD)

@@

reportlist <- list(runname=runname,
                   starttime=starttime,endtime=endtime,
                   zoneC=zoneC, zoneD=zoneD, product=product,
                   glb=glb,constants=constants
)

### 2020-06-10 data-raw/Tas_HS_PMs.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

Gradient1 is simply the proportional change in cpue from one year to the next. This implies that with a time-series of 28 years (as in the blockE13 data-set) we can obtain 27 comparisons. Formally, where $CE_{b,t}$ is the cpue in block/SMU $b$ in year $t$, then the Gradient1 for block/SMU $b$ is:

@@

In __aMSE__ we have the function _getgrad1_ to calculate the gradient. The function requires the input of a vector of cpue values (a minimum of two values) and it will output a vector of length one less than the input vector. If we input the vector _ab$cpue_ from teh blockE13 data-set we obtain Gradient1 values for 1993 - 2019, which have a range from -0.208 to 0.286, which is obviously asymmetric around the zero value.

>>>>

Gradient1 is simply the proportional change in cpue from one year to the next. This implies that with a time-series of 28 years (as in the blockE13 data-set) we can obtain 27 comparisons. Formally, where $CE_{b,t}$ is the cpue in block/SAU $b$ in year $t$, then the Gradient1 for block/SAU $b$ is:

@@

In __aMSE__ we have the function _getgrad1_ to calculate the gradient. The function requires the input of a vector of cpue values (a minimum of two values) and it will output a vector of length one less than the input vector. If we input the vector _ab$cpue_ from the blockE13 data-set we obtain Gradient1 values for 1993 - 2019, which have a range from -0.208 to 0.286, which is obviously asymmetric around the zero value.

### 2020-06-10 data-raw/development.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

* The region is simply to container for all the SMU (there may only be one). The notion is that the variation of population properties within an SMU can differ, but conditioning the populations is a whole subject for discussion in itself.

* Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular region, there is a need to include the option of having larval dispersal.

* Currently using a control.csv file to identify conditioning files, HCR files, and related. There is a region file containnig the strucutre of the simulated region, and a datafile containing the conditioning data.

* When constructing a region the constant parts are kept in regionC and the dynamic parts in regionD. Is the constant part always separate from the global part? Certainly not having to move such large objects around all the time should speed things along.

@@

* Currently using a spatial hierarchy region/SMU/population instead of zone/block/population, so this is now more general than only Tasmanian terminology. Stick to this hierarchy throughout the code. Consistency.

@@

* Once a particular run is constructed, but before the run proceeds, generate a series of plots and tables to characterize the simulated region.

@@

1) Initiate simulated region - use a ctrl file to define details of data-file, HCR function, and other initial conditions. May involve setting an initial depletion level.

2) conduct assessment of region status at start of each year (option for partial year?)

@@

* SMUname    - eg block1, block2
* popnum - population index   1, 2, 3, etc
* MaxDL      - maximum growth increment  eg 32
* sMaxDL     - variation of maxDL  eg  2.5
* L50        - Length at 50% MaxDL  eg 123
* sL50       - variation in L50  eg  5
* L50inc - L95 - L50 =  delta  eg 45
* sL50inc - variation of L50inc  eg 1.25
* SigMax - maximum variation around growth increment eg 4.6
* sSigMax - variation in SigMax  eg 0.1
* Wtb - weight-at-length exponent eg 3.161963
* sWtb - variation of Wtb  eg 0.148
* Wtbtoa - intercept power relationship betweeb Wtb and Wta 962.8098
* sWtbtoa - exponent of power relationship betweeb Wtb and Wta -14.3526
* Me - Natural mortality eg 0.2 maxage = 23, 0.1 maxage = 46 
* sMe - variation in Me  eg 0.003
* AvRec - log of unfished recruitment LnR0  eg 13
* sAvRec -   eg 1.0
* defsteep - Beverton-Holt Steepness  eg 0.6
* sdefsteep         eg 0.02
* L50C - length at 50% emergent  126.422
* sL50C                           1.5
* deltaC - length at increment 95% emergent 95 = 50 + deltaC eg  10
* sdeltaC                                      0.5
* MaxCEpar - maximum cpue tonnes/hour for defining catchability  0.37
* sMaxCEpar                              0.02
* selL50p - L50 of selectivity           0   (deviation from LML)
* selL95p - L95 of selectivity           1.5
* SaMa - maturity logistic a parameter  eg -16
* L50Mat - L50 for maturity b = -a/L50  eg  123.384
* sL50Mat                                   4.0
* larvalD - larval drift                    0.02  = 2%

2) A data-file containing details relating to the whole region

* numpop - number of populations
* PDFS   - The number of parameters with probability density functions
* minc   - minimum size class  2mm
* cw     - class width         2mm 
* Nclass - number of size classes 105
* nSMU   - number of Spatial Management Units  eg  2
* LML    - current Legal Minimum Length (LML, MLL, MLS, etc) e.g. 140
* randomseed - for repeatability of results - use a different one for each simulation
* Nyrs   - number of years for each simulation   eg 40
* firstYear - first year of simulation eg 1 (hypothetical) or 2014 (conditioned on fishery)
* fixYear - ?  eg 3  used in plotsimulation3
* projLML - the LML to be used in the simulated dynamics

@@

* label for run
* region file name
* datafile name
* HCR details file
* replicates
* HCRlabel
* output directory

@@

The unfished equilibrium is a classical starting point for any model, even if a decision is made to begin any simulations at some lower depletion level. We are using a one-year time-step but the dynamics occurs within each year. The dynamic part of the region has the following structure:

regionD - the dynamic part of hte regional structure

* SMU     : num [1:6] 1 1 2 2 2 2
* matureB : num [1:40, 1:6] 782 0 0 0 0 ...
* exploitB: num [1:40, 1:6] 892 0 0 0 0 ...
* catch   : num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* harvestR: num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* cpue    : num [1:40, 1:6] 399 0 0 0 0 ...
* recruit : num [1:40, 1:6] 630563 0 0 0 0 ...
* deplsB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
* depleB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
* catchN  : num [1:105, 1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* Nt      : num [1:105, 1:40, 1:6] 6.31e+05 2.34e-06 4.31e-05 ...

This represents the dynamics before any simulations have occurred, the unfished equilibrium state. Similarly, the constant part of the region, made up of a list for each population, has the following structure:

* Me      : num 0.198
* R0      : num 636147
* B0      : num 782
* effB0   : num 782
* ExB0    : num 892
* effExB0 : num 892
* MSY     : num 54.3
* MSYDepl : num 0.304
* bLML    : num 153
* popq    : num 0.000447
* SaM     : num 139
* popdef  : Named num [1:19] 34.45 122.75 164.92 4.61 -16 ....
* LML     : num [1:40] 132 132 132 132 132 132 132 132 132 132 ...
* G       : 'STM' num [1:105, 1:105] 2.12e-13 4.52e-12 8.33e-11 1.27e-09 1.62e-08 ...
* Maturity: num [1:105] 1.42e-07 1.78e-07 2.24e-07 2.83e-07 3.56e-07 ...
* WtL     : num [1:105] 0.000351 0.003438 0.013072 0.033719 0.070319 ...
* Emergent: num [1:105] 1.92e-16 3.44e-16 6.18e-16 1.11e-15 1.99e-15 ...
* Select  : num [1:105, 1:40] 2.87e-127 2.61e-125 2.37e-123 2.16e-121 1.96e-119 ...
* SelWt   : num [1:105, 1:40] 1.01e-130 8.97e-128 3.10e-125 7.28e-123 1.38e-120 ...
* MatWt   : num [1:105] 4.97e-11 6.13e-10 2.93e-09 9.53e-09 2.50e-08 ...
* SMU     : num 1

>>>>

* The zone is simply to container for all the SAU (there may only be one). The notion is that the variation of population properties within an SAU can differ, but conditioning the populations is a whole subject for discussion in itself.

* Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular zone, there is a need to include the option of having larval dispersal.

* Currently using a control.csv file to identify conditioning files, HCR files, and related. There is a zone file containnig the strucutre of the simulated zone, and a datafile containing the conditioning data.

* When constructing a zone the constant parts are kept in zoneC and the dynamic parts in zoneD. Is the constant part always separate from the global part? Certainly not having to move such large objects around all the time should speed things along.

@@

* Currently using a spatial hierarchy zone/SAU/population instead of zone/block/population, so this is now more general than only Tasmanian terminology. Stick to this hierarchy throughout the code. Consistency.

@@

* Once a particular run is constructed, but before the run proceeds, generate a series of plots and tables to characterize the simulated zone.

@@

1) Initiate simulated zone - use a ctrl file to define details of data-file, HCR function, and other initial conditions. May involve setting an initial depletion level.

2) conduct assessment of zone status at start of each year (option for partial year?)

@@

* SAUname    - eg block1, block2
* popnum - population index   1, 2, 3, etc
* MaxDL      - maximum growth increment  eg 32
* sMaxDL     - variation of maxDL  eg  2.5
* L50        - Length at 50% MaxDL  eg 123
* sL50       - variation in L50  eg  5
* L50inc - L95 - L50 =  delta  eg 45
* sL50inc - variation of L50inc  eg 1.25
* SigMax - maximum variation around growth increment eg 4.6
* sSigMax - variation in SigMax  eg 0.1
* Wtb - weight-at-length exponent eg 3.161963
* sWtb - variation of Wtb  eg 0.148
* Wtbtoa - intercept power relationship betweeb Wtb and Wta 962.8098
* sWtbtoa - exponent of power relationship betweeb Wtb and Wta -14.3526
* Me - Natural mortality eg 0.2 maxage = 23, 0.1 maxage = 46 
* sMe - variation in Me  eg 0.003
* AvRec - log of unfished recruitment LnR0  eg 13
* sAvRec -   eg 1.0
* defsteep - Beverton-Holt Steepness  eg 0.6
* sdefsteep         eg 0.02
* L50C - length at 50% emergent  126.422
* sL50C                           1.5
* deltaC - length at increment 95% emergent 95 = 50 + deltaC eg  10
* sdeltaC                                      0.5
* MaxCEpar - maximum cpue tonnes/hour for defining catchability  0.37
* sMaxCEpar                              0.02
* selL50p - L50 of selectivity           0   (deviation from LML)
* selL95p - L95 of selectivity           1.5
* SaMa - maturity logistic a parameter  eg -16
* L50Mat - L50 for maturity b = -a/L50  eg  123.384
* sL50Mat                                   4.0
* larvalD - larval drift                    0.02  = 2%

2) A data-file containing details relating to the whole zone

* numpop - number of populations
* PDFS   - The number of parameters with probability density functions
* minc   - minimum size class  2mm
* cw     - class width         2mm 
* Nclass - number of size classes 105
* nSAU   - number of Spatial Management Units  eg  2
* LML    - current Legal Minimum Length (LML, MLL, MLS, etc) e.g. 140
* randomseed - for repeatability of results - use a different one for each simulation
* Nyrs   - number of years for each simulation   eg 40
* firstYear - first year of simulation eg 1 (hypothetical) or 2014 (conditioned on fishery)
* fixYear - ?  eg 3  used in plotsimulation3
* projLML - the LML to be used in the simulated dynamics

@@

* label for run
* zone file name
* datafile name
* HCR details file
* replicates
* HCRlabel
* output directory

@@

The unfished equilibrium is a classical starting point for any model, even if a decision is made to begin any simulations at some lower depletion level. We are using a one-year time-step but the dynamics occurs within each year. The dynamic part of the zone has the following structure:

zoneD - the dynamic part of the zonal structure

* SAU     : num [1:6] 1 1 2 2 2 2
* matureB : num [1:40, 1:6] 782 0 0 0 0 ...
* exploitB: num [1:40, 1:6] 892 0 0 0 0 ...
* catch   : num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* harvestR: num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* cpue    : num [1:40, 1:6] 399 0 0 0 0 ...
* recruit : num [1:40, 1:6] 630563 0 0 0 0 ...
* deplsB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
* depleB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
* catchN  : num [1:105, 1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* Nt      : num [1:105, 1:40, 1:6] 6.31e+05 2.34e-06 4.31e-05 ...

This represents the dynamics before any simulations have occurred, the unfished equilibrium state. Similarly, the constant part of the zone, made up of a list for each population, has the following structure:

* Me      : num 0.198
* R0      : num 636147
* B0      : num 782
* effB0   : num 782
* ExB0    : num 892
* effExB0 : num 892
* MSY     : num 54.3
* MSYDepl : num 0.304
* bLML    : num 153
* popq    : num 0.000447
* SaM     : num 139
* popdef  : Named num [1:19] 34.45 122.75 164.92 4.61 -16 ....
* LML     : num [1:40] 132 132 132 132 132 132 132 132 132 132 ...
* G       : 'STM' num [1:105, 1:105] 2.12e-13 4.52e-12 8.33e-11 1.27e-09 1.62e-08 ...
* Maturity: num [1:105] 1.42e-07 1.78e-07 2.24e-07 2.83e-07 3.56e-07 ...
* WtL     : num [1:105] 0.000351 0.003438 0.013072 0.033719 0.070319 ...
* Emergent: num [1:105] 1.92e-16 3.44e-16 6.18e-16 1.11e-15 1.99e-15 ...
* Select  : num [1:105, 1:40] 2.87e-127 2.61e-125 2.37e-123 2.16e-121 1.96e-119 ...
* SelWt   : num [1:105, 1:40] 1.01e-130 8.97e-128 3.10e-125 7.28e-123 1.38e-120 ...
* MatWt   : num [1:105] 4.97e-11 6.13e-10 2.93e-09 9.53e-09 2.50e-08 ...
* SAU     : num 1

In the code (_setupzone_) the sequence is:

1. first generate zoneC from the constants and zone1. At that point each population structure contains everything except those things relating to productivity: B0, ExB0, MSY, MSYDepl and popq. makezoneC also updates glb by including the move matrix.
2. running the _makezone_ updates B0, ExB0, and popq in zoneC and generates the initial equilibrium for zoneD. As this is the unfished equilibrium the catch, harvestR, and catchN matrices are empty.
3. The final run of _moderegC_ internally runs _doproductivity_ which allows zoneC's MSY and MSYDepl entries to be updated for each population, as well as producing the product matrix.

In the zoneD matrices, the numbers-at-size are the numbers at the start of the year, similarly, the matureB is the mature biomass at the start of the year. The ExB0 is the exploitable biomass implied by the initial numbers-at-size, which means to calcualte it we need to apply half of natural mortality and growth before multiplyiung by selectivity and weight-at-length.
